A Virtual Research Environment to Document and Analyze Non-alphabetic Writing Systems. A Case Study for Maya Writing

### 2016-02-05 source/8/EAGLE_2016_submission_8_revised.tex
fixed biblio for maya

%all title info missing!
\title{A Virtual Research Environment to 
Document and Analyze Non-alphabetic Writing Systems. 
A Case Study for Maya Writing}
\author[BONN]{Katja Diederichs\corref{first}}
\author[BONN,MEL]{Sven Gronemeyer}
\author[BONN]{Christian Prager}
\author[BONN]{Elisabeth Wagner}
\author[GOT]{Franziska Diehr}
\author[GOT]{Maximilian Brodhun}
\author[BONN]{Nikolai Grube}
%
\address[GOT]{Niedersächsische Staats- und Universitätsbibliothek Göttingen}
\address[MEL]{La Trobe University Melbourne}
\address[BONN]{Rheinische Friedrich-Wilhelms-Universität Bonn}
%
\cortext[first]{Corresponding author. Email: diederichs@uni-bonn.de}

@@

\begin{enumerate}
\item Identification of hieroglyph blocks by alphanumeric classification 
\item Original spelling
\item Reading order of individual signs 
\item Number of signs within one block 
\item Identification and isolation of signs in one block 
\item Classification of signs based on the sign catalog of Eric Thompson \citep{Thompson1962}
\item Classification of signs in one block \citep{Thompson1962} 
\item Transliteration of individual signs 
\item Description of sign function (syllables, logographs)
\item Transliteration 
\item Broad transcription
\item Morphological segmentation 
\item Morphological analysis 
\item Determining congruence between block- and word-boundaries 
\item Determining syntactic function 
\item Determining sentence constituents 
\item Translation of single words 
\item Literal translation 
\item Rough translation
\end{enumerate}

@@

\subsection[Registering Object Data ]{Registering Object Data}
\noindent When designing the digital epigraphic work environment, one must also account for the current state of decipherment of
the script and language that are to be documented. Thus, Classic Mayan presents great challenges, as at least one third
of the known graphemes have not yet been completely deciphered. Even when a lexeme can be read, its etymology or
semantic domain is often unknown, and in some cases not even its lexical class can be determined, which causes
difficulties when attempting to read and understand a text. In order to decode an unknown sign, word, or sign sequence
in its respective context of use, for instance, semantic fields are generally investigated by means of applied
substitution. From a structuralist perspective, this method investigates the paradigmatic relation between two
linguistic entities as indicated by whether or not they appear in the same context \citep{Bussmann2002}. This step yields
results when sufficient quantities of data can be digitally compiled, marked up and investigated using corpus
linguistics.

@@

The decoding of Ugaritic cuneiform at the beginning of the twentieth century provides an example of a successful
decipherment that resulted from studying the relationship between object and textual data. Scholars suspected that the
word ``axe'' occurred in repeating, brief sign sequences on inscribed bronze axes. Building on this hypothesis, they
determined the phonetic values of individual signs using data on the Canaantic language and completely deciphered the
script \citep{Day2002}. Such investigations enable initial association of signs with both content and function in the
relevant language so that the signs may later be grammatically, morphologically, semantically, etc., defined and
eventually deciphered. This work also facilitates more precise typological classification of a writing system
\citep{Gronemeyer2015}.

@@

\subsubsection[Graphetics ]{Graphetics }
Graphetics concerns the formal structure of linguistic units and the structure of texts \citep{Crystal1997}. Classic Maya
hieroglyphic inscriptions, for instance, may display a very high degree of variation in writing styles for arranging
texts, a phenomenon which is deserving of investigation. In this manner, various systems of notation are studied in
their individual, social, and typographic aspects, for instance \citep{Bussmann2002}. Similarly, in paleography, script
decipherment is examined from graphetic perspective \citep{Bussmann2002}. The arrangement of texts, for instance, different
grapheme or block sizes of text fields, or different styles of hieroglyphic writing (equivalent to alphabetic font
style and size) are analyzed.

@@

\subsubsection{Graphemics }
When analyzing the contents of these texts, it is important to investigate meaningful variation underlying the manner in
which the writing is arranged and to thereby identify distinct units \citep{Bussmann2002}. Graphemics is devoted to this area
of research, and consequently to exploring the meaningfully distinct features of signs.

@@

\subsubsection[Signary: Sign Classification and Sign Catalogs ]{Signary: Sign Classification and Sign Catalogs }
\noindent The graphemic lexicon provides a central reference for marking up texts composed in non-alphabetic writing systems. In
the case of the Maya script, signs whose phonemic reading is unknown or unconfirmed are denoted using reference systems
from various classification catalogues \citep[e.g.]{Thompson1962}. Use of question marks for indicating unclear readings, for
example, should be avoided, because such signs impair machine readability and sometimes represent control characters or
part of an escape sequence. In this context, it should be determined what significance sign classifications and
nomenclatures have in their respective branches of epigraphic research. Also their role in transliterating and
transcribing texts in the respective writing system, as well as in sign classification is to be examined. Sign
inventories constitute a central authority in an epigraphic VRE, and they need to be modified or supplemented in
accordance with the ever-changing state of research as soon as newly discovered material is entered and signs are
identified that had not previously been isolated or identified as discrete graphemes.

>>>>

%all title info missing!
\title{A Virtual Research Environment to 
Document and Analyze Non-alphabetic Writing Systems. 
A Case Study for Maya Writing}
\author[BONN]{Katja Diederichs\corref{first}}
\author[BONN,MEL]{Sven Gronemeyer}
\author[BONN]{Christian Prager}
\author[BONN]{Elisabeth Wagner}
\author[GOT]{Franziska Diehr}
\author[GOT]{Maximilian Brodhun}
\author[BONN]{Nikolai Grube}
%

\address[BONN]{Rheinische Friedrich-Wilhelms-Universität Bonn}
\address[MEL]{La Trobe University Melbourne}
\address[GOT]{Niedersächsische Staats- und Universitätsbibliothek Göttingen}

%
\cortext[first]{Corresponding author. Email: diederichs@uni-bonn.de}

@@

\begin{enumerate}
\item Identification of hieroglyph blocks by alphanumeric classification 
\item Original spelling
\item Reading order of individual signs 
\item Number of signs within one block 
\item Identification and isolation of signs in one block 
\item Classification of signs based on the sign catalog Eric Thompson
\item Classification of signs in one block 
\item Transliteration of individual signs 
\item Description of sign function (syllables, logographs)
\item Transliteration 
\item Broad transcription
\item Morphological segmentation 
\item Morphological analysis 
\item Determining congruence between block- and word-boundaries 
\item Determining syntactic function

\end{enumerate}

@@

\subsection[Registering Object Data ]{Registering Object Data}
\noindent When designing the digital epigraphic work environment, one must also account for the current state of decipherment of
the script and language that are to be documented. Thus, Classic Mayan presents great challenges, as at least one third
of the known graphemes have not yet been completely deciphered. Even when a lexeme can be read, its etymology or
semantic domain is often unknown, and in some cases not even its lexical class can be determined, which causes
difficulties when attempting to read and understand a text. In order to decode an unknown sign, word, or sign sequence
in its respective context of use, for instance, semantic fields are generally investigated by means of applied
substitution. From a structuralist perspective, this method investigates the paradigmatic relation between two
linguistic entities as indicated by whether or not they appear in the same context. 
%\citep{Bussmann2002} 
This step yields
results when sufficient quantities of data can be digitally compiled, marked up and investigated using corpus
linguistics.

@@

The decoding of Ugaritic cuneiform at the beginning of the twentieth century provides an example of a successful
decipherment that resulted from studying the relationship between object and textual data. Scholars suspected that the
word ``axe'' occurred in repeating, brief sign sequences on inscribed bronze axes. Building on this hypothesis, they
determined the phonetic values of individual signs using data on the %Canaantic 
language and completely deciphered the script \citep{Day2002}. Such investigations enable initial association of signs with both content and function in the
relevant language so that the signs may later be grammatically, morphologically, semantically, etc., defined and
eventually deciphered. This work also facilitates more precise typological classification of a writing system
\citep{Gronemeyer2015}.

@@

\subsubsection[Graphetics ]{Graphetics }
Graphetics concerns the formal structure of linguistic units and the structure of texts \citep{Crystal1997}. Classic Maya
hieroglyphic inscriptions, for instance, may display a very high degree of variation in writing styles for arranging
texts, a phenomenon which is deserving of investigation. In this manner, various systems of notation are studied in
their individual, social, and typographic aspects, for instance.% \citep{Bussmann2002}. 
Similarly, in paleography, script
decipherment is examined from graphetic perspective. %\citep{Bussmann2002}. 
The arrangement of texts, for instance, different
grapheme or block sizes of text fields, or different styles of hieroglyphic writing (equivalent to alphabetic font
style and size) are analyzed.

@@

\subsubsection{Graphemics }
When analyzing the contents of these texts, it is important to investigate meaningful variation underlying the manner in
which the writing is arranged and to thereby identify distinct units. 
%\citep{Bussmann2002}. 
Graphemics is devoted to this area
of research, and consequently to exploring the meaningfully distinct features of signs.

@@

\subsubsection[Signary: Sign Classification and Sign Catalogs ]{Signary: Sign Classification and Sign Catalogs }
\noindent The graphemic lexicon provides a central reference for marking up texts composed in non-alphabetic writing systems. In
the case of the Maya script, signs whose phonemic reading is unknown or unconfirmed are denoted using reference systems
from various classification catalogues. 
%\citep[e.g.]{Thompson1962}. 
Use of question marks for indicating unclear readings, for
example, should be avoided, because such signs impair machine readability and sometimes represent control characters or
part of an escape sequence. In this context, it should be determined what significance sign classifications and
nomenclatures have in their respective branches of epigraphic research. Also their role in transliterating and
transcribing texts in the respective writing system, as well as in sign classification is to be examined. Sign
inventories constitute a central authority in an epigraphic VRE, and they need to be modified or supplemented in
accordance with the ever-changing state of research as soon as newly discovered material is entered and signs are
identified that had not previously been isolated or identified as discrete graphemes.
